Recent demography drives changes in linked selection across the maize genome

### 2015-11-25 NatGen/manuscript.tex
updates

\begin{abstract}
  \noindent \bf
  \noindent
Genetic diversity is shaped by the interaction of drift and selection, but the details of this interaction are not well understood. Although a population's propensity to drift is typically summarized by its long-term effective population size ($N_e$ ), rapidly changing population demographics complicate this relationship. To better understand how changing demography impacts selection, we used whole-genome sequencing data to investigate patterns of linked selection in domesticated and wild maize (teosinte). We produce the first whole-genome estimate of the demography of maize domestication, showing that maize was reduced to approximately 5\% the population size of teosinte before it experienced rapid expansion post-domestication to population sizes much larger than its ancestor. Evaluation of patterns of nucleotide diversity in and near genes shows little evidence of selection on beneficial amino acid substitutions, and that the domestication bottleneck lead to a decline in the efficiency of purifying selection in maize.  Young alleles, however, show evidence of much stronger purifying selection in maize, reflecting the much larger effective size of present day populations. Our results demonstrate that recent demographic change --- a hallmark of many species including both humans and crops ----  can have immediate and wide-ranging impacts on diversity that conflict with would-be expectations based on $N_e$ alone.
\end{abstract}

@@

\section*{AUTHOR CONTRIBUTIONS}
TMB and JRI devised this study. TB, LW, and KC analyzed the data. AD performed early-stage simulations. MBH provided advice. TB, JRI, and MBH wrote the manuscript.

>>>>

\begin{abstract}
  \noindent \bf
  \noindent
Genetic diversity is shaped by the interaction of drift and selection, but the details of this interaction are not well understood. Although a population's propensity to drift is typically summarized by its long-term effective population size ($N_e$ ), rapidly changing population demographics complicate this relationship. To better understand how changing demography impacts selection, we used whole-genome sequencing data to investigate patterns of linked selection in domesticated and wild maize (teosinte). We produce the first whole-genome estimate of the demography of maize domestication, showing that maize was reduced to approximately 5\% the population size of teosinte before it experienced rapid expansion post-domestication to population sizes much larger than its ancestor. Evaluation of patterns of nucleotide diversity in and near genes shows little evidence of selection on beneficial amino acid substitutions, and that the domestication bottleneck lead to a decline in the efficiency of purifying selection in maize.  Young alleles, however, show evidence of much stronger purifying selection in maize, reflecting the much larger effective size of present day populations. Our results demonstrate that recent demographic change --- a hallmark of many species including both humans and crops ---  can have immediate and wide-ranging impacts on diversity that conflict with would-be expectations based on $N_e$ alone.
\end{abstract}

@@

\section*{AUTHOR CONTRIBUTIONS}
TMB and JRI devised this study. TMB, LW, and KC analyzed the data. AD performed early-stage simulations. MBH provided advice. TMB, JRI, and MBH wrote the manuscript.
